---
title: "DataSci 306 Final Project"
author: "Haohong Zheng & Ziqing Li"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
```

## Investigating the Internet Movie Database (IMDB)

## Load Preprocessed Data



```{r load-data}

title_basics <- read_rds("data/title_basics_sample.rda")
title_ratings <- read_rds("data/title_ratings_sample.rda")
name_basics <- read_rds("data/name_basics_sample.rda") |>
  select(nconst, primaryName, birthYear, deathYear, primaryProfession, knownForTitles)
title_principals <- read_rds("data/title_principals_sample.rda") |>
  select(tconst, ordering, nconst, category, job)
```

---

## Part II: EDA of Individual Tables


### Table 1: title_basics

```{r eda-tb-structure}
glimpse(title_basics)
```

```{r eda-tb-numeric}
title_basics |>
  select(where(is.numeric)) |>
  summary()
```

```{r eda-tb-categorical}
title_basics |>
  count(titleType, sort = TRUE)
```

```{r eda-tb-plot}
title_basics |>
  filter(!is.na(runtimeMinutes), runtimeMinutes < 300) |>
  filter(titleType %in% c("movie", "tvSeries", "tvMovie", "short", "tvMiniSeries")) |>
  ggplot(aes(x = titleType, y = runtimeMinutes, fill = titleType)) +
  geom_violin(alpha = 0.7) +
  geom_boxplot(width = 0.1, alpha = 0.5) +
  labs(title = "Distribution of Runtime by Title Type",
       x = "Title Type", y = "Runtime (minutes)") +
  theme_minimal() +
  theme(legend.position = "none") +
  coord_flip()
```

#### How runtimeMinutes varies by two grouping variables


```{r runtime-group1}
title_basics |>
  filter(!is.na(runtimeMinutes)) |>
  group_by(titleType) |>
  summarize(
    mean_runtime = mean(runtimeMinutes, na.rm = TRUE),
    median_runtime = median(runtimeMinutes, na.rm = TRUE),
    sd_runtime = sd(runtimeMinutes, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) |>
  arrange(desc(mean_runtime))
```

```{r runtime-group2}
title_basics |>
  filter(!is.na(runtimeMinutes), !is.na(startYear)) |>
  mutate(decade = (startYear %/% 10) * 10) |>
  group_by(decade) |>
  summarize(
    mean_runtime = mean(runtimeMinutes, na.rm = TRUE),
    median_runtime = median(runtimeMinutes, na.rm = TRUE),
    sd_runtime = sd(runtimeMinutes, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) |>
  arrange(decade)
```

#### How many titles have a different known name vs original name?


```{r diff-names}
diff_name_count <- title_basics |>
  filter(!is.na(primaryTitle), !is.na(originalTitle)) |>
  mutate(name_differs = primaryTitle != originalTitle) |>
  count(name_differs)

diff_name_count

n_diff <- diff_name_count |> filter(name_differs == TRUE) |> pull(n)
cat("Number of titles with different known vs original name:", n_diff, "\n")
```

#### Conditional distribution of release year by name difference

```{r year-by-name-diff}
title_basics |>
  filter(!is.na(primaryTitle), !is.na(originalTitle), !is.na(startYear)) |>
  mutate(name_differs = ifelse(primaryTitle != originalTitle,
                               "Different Name", "Same Name")) |>
  ggplot(aes(x = startYear, fill = name_differs)) +
  geom_density(alpha = 0.5) +
  labs(title = "Distribution of Release Year by Name Difference",
       x = "Release Year", y = "Density", fill = "Name Status") +
  theme_minimal() +
  annotate("text", x = 1950, y = 0.02,
           label = "Titles with different names\nare more common in recent years",
           size = 3.5, hjust = 0)
```

Titles with different known names vs original names tend to be more common in recent decades, likely reflecting increased international distribution and localization of content.

---

### Table 2: title_ratings

```{r eda-tr-structure}
glimpse(title_ratings)

title_ratings |>
  select(where(is.numeric)) |>
  summary()
```

```{r eda-tr-plot}
title_ratings |>
  ggplot(aes(x = numVotes, y = averageRating)) +
  geom_point(alpha = 0.1, size = 0.5, color = "steelblue") +
  stat_smooth(method = "loess", color = "red", se = TRUE) +
  scale_x_log10(labels = scales::comma) +
  labs(title = "Average Rating vs Number of Votes",
       x = "Number of Votes (log scale)", y = "Average Rating") +
  theme_minimal()
```

#### Rating groups: higher vs lower rated


```{r rating-groups}
title_ratings |>
  mutate(rating_group = cut(averageRating,
                            breaks = c(0, 4, 7, 10),
                            labels = c("Low (0-4)", "Medium (4-7)", "High (7-10)"),
                            include.lowest = TRUE)) |>
  group_by(rating_group) |>
  summarize(
    count = n(),
    mean_votes = mean(numVotes, na.rm = TRUE),
    median_votes = median(numVotes, na.rm = TRUE),
    total_votes = sum(numVotes, na.rm = TRUE),
    .groups = "drop"
  )
```

Higher rated titles tend to have more votes on average, suggesting that popular and well-known titles attract both more viewers and higher ratings.

---

### Table 3: name_basics

```{r eda-nb-structure}
glimpse(name_basics)

name_basics |>
  select(where(is.numeric)) |>
  summary()
```

```{r eda-nb-categorical}
name_basics |>
  separate_rows(primaryProfession, sep = ",") |>
  count(primaryProfession, sort = TRUE) |>
  head(15)
```

```{r eda-nb-plot}
name_basics |>
  separate_rows(primaryProfession, sep = ",") |>
  count(primaryProfession, sort = TRUE) |>
  head(10) |>
  ggplot(aes(x = reorder(primaryProfession, n), y = n, fill = primaryProfession)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(title = "Top 10 Primary Professions in IMDB",
       x = "Profession", y = "Count") +
  theme_minimal() +
  scale_fill_viridis_d()
```

#### Number of titles each person is known for


```{r known-for-count}
name_basics |>
  filter(!is.na(knownForTitles)) |>
  mutate(n_known_for = str_count(knownForTitles, ",") + 1) |>
  ggplot(aes(x = n_known_for)) +
  geom_bar(fill = "steelblue") +
  labs(title = "Distribution of Number of Titles Each Person is Known For",
       x = "Number of Known-For Titles", y = "Count") +
  theme_minimal()
```

#### Age of cast members


```{r age-deceased}
name_basics_age <- name_basics |>
  filter(!is.na(birthYear)) |>
  mutate(
    is_deceased = !is.na(deathYear),
    age = ifelse(is_deceased,
                 deathYear - birthYear,
                 2026 - birthYear)
  ) |>
  filter(age > 0, age < 120)


name_basics_age |>
  filter(is_deceased) |>
  ggplot(aes(x = age)) +
  geom_histogram(binwidth = 5, fill = "coral", alpha = 0.7, color = "white") +
  labs(title = "Age Distribution of Deceased Cast Members",
       x = "Age at Death", y = "Count") +
  theme_minimal()
```

```{r age-living}
name_basics_age |>
  filter(!is_deceased) |>
  ggplot(aes(x = age)) +
  geom_histogram(binwidth = 5, fill = "skyblue", alpha = 0.7, color = "white") +
  labs(title = "Age Distribution of Living Cast Members",
       x = "Current Age", y = "Count") +
  theme_minimal()
```

#### Actors named Tom/Thomas/Thom/Tomas

```{r tom-actors}
tom_actors <- name_basics |>
  filter(str_detect(primaryName, "^(Tom|Thomas|Thom|Tomas)\\b"))

cat("Number of actors with first name Tom/Thomas/Thom/Tomas:", nrow(tom_actors), "\n")
```

---

### Table 4: title_principals

```{r eda-tp-structure}
glimpse(title_principals)

title_principals |>
  count(category, sort = TRUE)
```

```{r eda-tp-plot}
title_principals |>
  filter(category %in% c("actor", "actress", "director", "writer", "producer")) |>
  group_by(category, ordering) |>
  summarize(count = n(), .groups = "drop") |>
  filter(ordering <= 10) |>
  ggplot(aes(x = ordering, y = category, fill = count)) +
  geom_tile() +
  scale_fill_viridis_c() +
  labs(title = "Heatmap of Ordering Position by Category",
       x = "Ordering", y = "Category", fill = "Count") +
  theme_minimal()
```

#### Alliterative titles


```{r alliteration}
alliterative <- title_basics |>
  filter(!is.na(primaryTitle)) |>
  mutate(
    words = str_extract_all(primaryTitle, "\\b[A-Za-z]"),
    first_letters = map_chr(words, ~ {
      letters_vec <- toupper(.x)
      if (length(letters_vec) <= 1) return("skip")  
      if (all(letters_vec == letters_vec[1])) return("alliterative")
      return("not")
    })
  ) |>
  filter(first_letters == "alliterative")

cat("Number of alliterative titles:", nrow(alliterative), "\n")

alliterative |>
  select(primaryTitle) |>
  head(10)
```

---

## Part III: Pivoting

### One row per title-genre combination


```{r genres-tall}
titles_genres <- title_basics |>
  filter(!is.na(genres)) |>
  separate_rows(genres, sep = ",")

cat("Original rows:", nrow(title_basics), "\n")
cat("After separating genres:", nrow(titles_genres), "\n")
head(titles_genres)
```

### Genre counts over time (line plot)

```{r genres-over-time}
top_genres <- titles_genres |>
  count(genres, sort = TRUE) |>
  head(8) |>
  pull(genres)

titles_genres |>
  filter(!is.na(startYear), genres %in% top_genres, startYear >= 1920) |>
  count(startYear, genres) |>
  ggplot(aes(x = startYear, y = n, color = genres)) +
  geom_line(linewidth = 0.8) +
  labs(title = "Count of Top Genres Over Time",
       x = "Year", y = "Count", color = "Genre") +
  theme_minimal()
```

### Genre correlation using model.matrix



```{r genre-wide}
# 先弄一个 genre 作为 factor
titles_genre_factor <- titles_genres |>
  select(tconst, genres) |>
  distinct() |>
  mutate(genres = factor(genres))

# model.matrix 建 one-hot 编码
genre_matrix <- model.matrix(~ genres - 1, data = titles_genre_factor)

genre_wide <- as_tibble(genre_matrix) |>
  mutate(tconst = titles_genre_factor$tconst) |>
  group_by(tconst) |>
  summarize(across(starts_with("genres"), max), .groups = "drop")

genre_cols <- genre_wide |> select(starts_with("genres"))
genre_cor <- cor(genre_cols)

genre_cor[upper.tri(genre_cor, diag = TRUE)] <- NA
max_idx <- which(genre_cor == max(genre_cor, na.rm = TRUE), arr.ind = TRUE)

cat("Most correlated genre pair:",
    rownames(genre_cor)[max_idx[1, 1]], "&",
    colnames(genre_cor)[max_idx[1, 2]],
    "\nCorrelation:", genre_cor[max_idx[1, 1], max_idx[1, 2]], "\n")
```

---

## Part IV: Joining Tables

### Genre + Ratings: highest and lowest rated genre

```{r genre-ratings}
genres_ratings <- titles_genres |>
  inner_join(title_ratings, by = "tconst")

genre_rating_summary <- genres_ratings |>
  group_by(genres) |>
  summarize(
    mean_rating = mean(averageRating, na.rm = TRUE),
    median_rating = median(averageRating, na.rm = TRUE),
    count = n(),
    .groups = "drop"
  ) |>
  arrange(desc(mean_rating))

genre_rating_summary

cat("Highest rated genre:", genre_rating_summary |> slice(1) |> pull(genres), "\n")
cat("Lowest rated genre:", genre_rating_summary |> slice(n()) |> pull(genres), "\n")
```

### Stacked bar chart: genre proportions over time

```{r genre-proportions}
genres_ratings |>
  filter(!is.na(startYear), genres %in% top_genres) |>
  mutate(decade = (startYear %/% 10) * 10) |>
  filter(decade >= 1950) |>
  count(decade, genres) |>
  group_by(decade) |>
  mutate(prop = n / sum(n)) |>
  ungroup() |>
  mutate(genres = fct_reorder(genres, prop, .fun = sum, .desc = TRUE)) |>
  ggplot(aes(x = factor(decade), y = prop, fill = genres)) +
  geom_col(position = "stack") +
  labs(title = "Proportions of Genres Over Decades",
       x = "Decade", y = "Proportion", fill = "Genre") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Some genres like Drama remain consistently dominant. Documentary has increased in proportion over recent decades, while some traditional genres have slightly decreased.

### Number of ratings over time

```{r ratings-over-time}
basics_ratings <- title_basics |>
  inner_join(title_ratings, by = "tconst")

basics_ratings |>
  filter(!is.na(startYear), startYear >= 1920) |>
  group_by(startYear) |>
  summarize(
    total_votes = sum(numVotes, na.rm = TRUE),
    mean_votes = mean(numVotes, na.rm = TRUE),
    n_titles = n(),
    .groups = "drop"
  ) |>
  ggplot(aes(x = startYear, y = total_votes)) +
  geom_line(color = "steelblue", linewidth = 0.8) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Total Number of Ratings Over Time",
       x = "Release Year", y = "Total Votes") +
  theme_minimal()
```

```{r ratings-numerical}
basics_ratings |>
  filter(!is.na(startYear)) |>
  mutate(era = case_when(
    startYear < 1970 ~ "Before 1970",
    startYear < 2000 ~ "1970-1999",
    TRUE ~ "2000+"
  )) |>
  group_by(era) |>
  summarize(
    mean_votes_per_title = mean(numVotes, na.rm = TRUE),
    median_votes_per_title = median(numVotes, na.rm = TRUE),
    total_titles = n(),
    .groups = "drop"
  )
```

The number of ratings has increased dramatically over time, especially for titles released after 2000, reflecting the growth of online rating platforms.

### Top 10 people by median rating

```{r top10-people}
people_ratings <- title_principals |>
  inner_join(title_ratings, by = "tconst") |>
  inner_join(name_basics |> select(nconst, primaryName), by = "nconst")

top10 <- people_ratings |>
  group_by(nconst, primaryName) |>
  summarize(
    median_rating = median(averageRating, na.rm = TRUE),
    n_titles = n(),
    .groups = "drop"
  ) |>
  filter(n_titles >= 3) |> 
  arrange(desc(median_rating)) |>
  head(10)

top10
```

#### Genre proportions for top 10 rated principals

```{r top10-genres}
top10_titles <- title_principals |>
  filter(nconst %in% top10$nconst) |>
  inner_join(titles_genres |> select(tconst, genres), by = "tconst")

top10_titles |>
  count(genres, sort = TRUE) |>
  mutate(prop = n / sum(n)) |>
  head(10)
```

#### Ratings vs years trend

```{r ratings-vs-years}
basics_ratings |>
  filter(!is.na(startYear), startYear >= 1920) |>
  group_by(startYear) |>
  summarize(mean_rating = mean(averageRating, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(x = startYear, y = mean_rating)) +
  geom_point(alpha = 0.5, color = "steelblue") +
  geom_smooth(method = "loess", color = "red") +
  labs(title = "Average Rating Over Years",
       x = "Release Year", y = "Mean Rating") +
  theme_minimal()
```

Ratings tend to be higher for older titles, likely due to survivorship bias — only well-regarded older titles remain widely known and rated.

### Highest median known-for rating


```{r known-for-rating}
known_for_ratings <- name_basics |>
  filter(!is.na(knownForTitles)) |>
  select(nconst, primaryName, knownForTitles) |>
  separate_rows(knownForTitles, sep = ",") |>
  rename(tconst = knownForTitles) |>
  inner_join(title_ratings, by = "tconst")

known_for_ratings |>
  group_by(nconst, primaryName) |>
  summarize(
    median_kf_rating = median(averageRating, na.rm = TRUE),
    n_known_for = n(),
    .groups = "drop"
  ) |>
  filter(n_known_for >= 2) |>
  arrange(desc(median_kf_rating)) |>
  head(10)
```

---

## Part V: Profiling and Parallel Processing

### Profiling


```{r profiling}
prof_time <- system.time({
  full_result <- title_principals |>
    inner_join(title_ratings, by = "tconst") |>
    inner_join(title_basics |> select(tconst, primaryTitle, startYear), by = "tconst") |>
    group_by(category) |>
    summarize(
      mean_rating = mean(averageRating, na.rm = TRUE),
      n = n(),
      .groups = "drop"
    )
})

cat("Time for triple join + grouping:", prof_time["elapsed"], "seconds\n")
full_result
```

The join operations are the most expensive steps. Selecting only the necessary columns before joining (e.g., `select(tconst, primaryTitle, startYear)`) reduces memory usage and speeds up the computation significantly compared to joining entire tables.

### Parallelization

```{r parallel}
library(parallel)

genres_subset <- titles_genres |> filter(genres %in% top_genres)

seq_time <- system.time({
  seq_result <- genres_subset |>
    group_by(genres) |>
    summarize(
      n = n(),
      mean_year = mean(startYear, na.rm = TRUE),
      .groups = "drop"
    )
})

n_cores <- max(1, detectCores() - 1)

par_time <- system.time({
  genre_list <- split(genres_subset, genres_subset$genres)
  
  par_result <- mclapply(genre_list, function(df) {
    tibble(
      genres = df$genres[1],
      n = nrow(df),
      mean_year = mean(df$startYear, na.rm = TRUE)
    )
  }, mc.cores = n_cores) |>
    bind_rows()
})

cat("Sequential time:", seq_time["elapsed"], "seconds\n")
cat("Parallel time:", par_time["elapsed"], "seconds\n")
```

### Benchmarking: two ways to count comma-separated items

```{r benchmarking}
count_items_v1 <- function(x) {
  str_count(x, ",") + 1
}

count_items_v2 <- function(x) {
  lengths(strsplit(x, ",", fixed = TRUE))
}

test_strings <- name_basics$knownForTitles[!is.na(name_basics$knownForTitles)]

# bench mark
bench_result <- bench::mark(
  str_count = count_items_v1(test_strings),
  strsplit = count_items_v2(test_strings),
  check = FALSE,
  min_iterations = 5
)

bench_result |> select(expression, median, mem_alloc)
```

The `str_count` method is generally faster and more memory efficient because it avoids creating intermediate list objects. We recommend using `str_count` for counting items in comma-separated strings.

---

## Part VI: Shiny Applications

### Application 1: Interactive IMDB Explorer


```{r shiny-app1, eval=FALSE}
library(shiny)

ui1 <- fluidPage(
  titlePanel("IMDB Explorer"),
  sidebarLayout(
    sidebarPanel(
      selectInput("genre", "Select Genre:",
                  choices = sort(unique(titles_genres$genres)),
                  selected = "Drama"),
      sliderInput("year_range", "Year Range:",
                  min = 1900, max = 2025,
                  value = c(1990, 2025)),
      numericInput("min_votes", "Minimum Votes:",
                   value = 100, min = 1, max = 100000)
    ),
    mainPanel(
      plotOutput("rating_plot"),
      plotOutput("runtime_plot"),
      tableOutput("top_titles")
    )
  )
)

server1 <- function(input, output) {
  
  filtered_data <- reactive({
    titles_genres |>
      filter(genres == input$genre,
             !is.na(startYear),
             startYear >= input$year_range[1],
             startYear <= input$year_range[2]) |>
      inner_join(title_ratings, by = "tconst") |>
      filter(numVotes >= input$min_votes)
  })
  
  output$rating_plot <- renderPlot({
    filtered_data() |>
      ggplot(aes(x = startYear, y = averageRating)) +
      geom_point(alpha = 0.3, color = "steelblue") +
      geom_smooth(method = "loess", color = "red") +
      labs(title = paste("Ratings for", input$genre, "Over Time"),
           x = "Year", y = "Average Rating") +
      theme_minimal()
  })
  
  output$runtime_plot <- renderPlot({
    filtered_data() |>
      filter(!is.na(runtimeMinutes), runtimeMinutes < 300) |>
      ggplot(aes(x = runtimeMinutes)) +
      geom_histogram(binwidth = 10, fill = "coral", alpha = 0.7) +
      labs(title = paste("Runtime Distribution for", input$genre),
           x = "Runtime (minutes)", y = "Count") +
      theme_minimal()
  })
  
  output$top_titles <- renderTable({
    filtered_data() |>
      arrange(desc(averageRating)) |>
      select(primaryTitle, startYear, averageRating, numVotes) |>
      head(10)
  })
}

shinyApp(ui1, server1)
```

### Application 2: Category-Based Job Explorer


```{r shiny-app2, eval=FALSE}
ui2 <- fluidPage(
  titlePanel("IMDB Principals Explorer"),
  sidebarLayout(
    sidebarPanel(
      selectInput("category", "Select Category:",
                  choices = sort(unique(title_principals$category))),
      uiOutput("job_selector")
    ),
    mainPanel(
      plotOutput("category_plot"),
      tableOutput("category_table")
    )
  )
)

server2 <- function(input, output) {
  
  output$job_selector <- renderUI({
    jobs <- title_principals |>
      filter(category == input$category, !is.na(job)) |>
      count(job, sort = TRUE) |>
      head(20) |>
      pull(job)
    
    selectInput("job", "Select Specific Job:",
                choices = c("All", jobs),
                selected = "All")
  })
  
  filtered <- reactive({
    df <- title_principals |>
      filter(category == input$category)
    
    if (!is.null(input$job) && input$job != "All") {
      df <- df |> filter(job == input$job)
    }
    
    df |>
      inner_join(title_basics |> select(tconst, primaryTitle, startYear, genres),
                 by = "tconst") |>
      inner_join(title_ratings, by = "tconst")
  })
  
  output$category_plot <- renderPlot({
    filtered() |>
      filter(!is.na(startYear)) |>
      group_by(startYear) |>
      summarize(mean_rating = mean(averageRating, na.rm = TRUE), .groups = "drop") |>
      ggplot(aes(x = startYear, y = mean_rating)) +
      geom_line(color = "steelblue") +
      geom_smooth(method = "loess", color = "red", se = FALSE) +
      labs(title = paste("Average Rating Over Time for", input$category),
           x = "Year", y = "Average Rating") +
      theme_minimal()
  })
  
  output$category_table <- renderTable({
    filtered() |>
      arrange(desc(averageRating)) |>
      select(primaryTitle, startYear, averageRating, numVotes) |>
      distinct() |>
      head(15)
  })
}

shinyApp(ui2, server2)
```

---

## Extra Credit: Six Degrees of Kevin Bacon

```{r kevin-bacon, eval=FALSE}
ui_kb <- fluidPage(
  titlePanel("Six Degrees of Kevin Bacon"),
  sidebarLayout(
    sidebarPanel(
      textInput("title_search", "Enter a Movie/TV Title:", value = ""),
      actionButton("search_btn", "Search"),
      hr(),
      uiOutput("cast_selector"),
      actionButton("select_btn", "Select Person"),
      hr(),
      textOutput("step_counter"),
      textOutput("game_status")
    ),
    mainPanel(
      tableOutput("cast_list"),
      h4("Connection History"),
      tableOutput("connection_history")
    )
  )
)

server_kb <- function(input, output, session) {
  
  game_state <- reactiveValues(
    steps = 0,
    history = data.frame(Step = integer(), Person = character(),
                         stringsAsFactors = FALSE),
    current_cast = NULL,
    won = FALSE,
    lost = FALSE
  )
  

  observeEvent(input$search_btn, {
    req(input$title_search)
    
    matching_titles <- title_basics |>
      filter(str_detect(primaryTitle,
                        fixed(input$title_search, ignore_case = TRUE))) |>
      head(5)
    
    if (nrow(matching_titles) > 0) {
      cast <- title_principals |>
        filter(tconst %in% matching_titles$tconst) |>
        inner_join(name_basics |> select(nconst, primaryName), by = "nconst") |>
        select(nconst, primaryName, tconst) |>
        inner_join(matching_titles |> select(tconst, primaryTitle), by = "tconst") |>
        distinct()
      
      game_state$current_cast <- cast
      
      if ("nm0000102" %in% cast$nconst) {
        game_state$won <- TRUE
      }
    }
  })
  
  output$cast_selector <- renderUI({
    req(game_state$current_cast)
    selectInput("person", "Select a Person:",
                choices = setNames(game_state$current_cast$nconst,
                                   game_state$current_cast$primaryName))
  })
  
  observeEvent(input$select_btn, {
    req(input$person)
    if (game_state$won || game_state$lost) return()
    
    game_state$steps <- game_state$steps + 1
    
    person_name <- game_state$current_cast |>
      filter(nconst == input$person) |>
      pull(primaryName) |>
      first()
    
    game_state$history <- rbind(
      game_state$history,
      data.frame(Step = game_state$steps, Person = person_name,
                 stringsAsFactors = FALSE)
    )
    
    if (game_state$steps >= 6) {
      game_state$lost <- TRUE
      return()
    }

    person_titles <- title_principals |>
      filter(nconst == input$person) |>
      pull(tconst)

    new_cast <- title_principals |>
      filter(tconst %in% person_titles) |>
      inner_join(name_basics |> select(nconst, primaryName), by = "nconst") |>
      inner_join(title_basics |> select(tconst, primaryTitle), by = "tconst") |>
      select(nconst, primaryName, tconst, primaryTitle) |>
      distinct()
    
    game_state$current_cast <- new_cast
    
    if ("nm0000102" %in% new_cast$nconst) {
      game_state$won <- TRUE
    }
  })
  
  output$cast_list <- renderTable({
    req(game_state$current_cast)
    game_state$current_cast |>
      select(primaryName, primaryTitle) |>
      distinct() |>
      head(50)
  })
  
  output$step_counter <- renderText({
    paste("Steps taken:", game_state$steps, "/ 6")
  })
  
  output$game_status <- renderText({
    if (game_state$won) return("YOU WON! Kevin Bacon found!")
    if (game_state$lost) return("GAME OVER! You exceeded 6 steps.")
    "Keep searching..."
  })
  
  output$connection_history <- renderTable({
    game_state$history
  })
}

shinyApp(ui_kb, server_kb)
```
